---
title: "iPSC Cardiomyocyte Characterization (Axol and CDI)"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmdformats::html_clean:
    highlight: kate
    toc_depth: 1
    self_contained: true
    gallery: true
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)
library(svglite)

optisvg = function(before, options, envir) {
  if (before) return()
  if (!file.exists(fig_path("svg")[1])) return()
  for (f in fig_path("svg")) {
   system2('svgo', shQuote(f))
  }
}

optipng = function(before, options, envir) {
  if (before) return()
  if (!file.exists(fig_path("png")[1])) return()
  for (f in fig_path("png")) {
   system2('mogrify', c('-trim', '-bordercolor white', '-border 10', shQuote(f)))
   system2('pngquant', c('--speed 10', '--nofs', '-f', '--ext .png', shQuote(f)))
  }
}

## Global options
options(max.print="95")
opts_chunk$set(echo=FALSE,
               cache=FALSE,
               prompt=FALSE,
               tidy=FALSE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               fig.retina = 1,
               #dpi = 72,
               dev = 'svg',
               opt.svg = TRUE)
               #opt.png = TRUE)
opts_knit$set(width=95)

## IMG optimization
knit_hooks$set(opt.svg = optisvg)
#knit_hooks$set(opt.png = optipng)
```

```{r LoadLibraries, results='hide'}
# Always Load these libraries
# CRAN
library(assertthat)
library(dplyr)
library(tidyr)
library(reshape2)
library(openxlsx)
library(DT)
library(pheatmap)

# Bioconductor
library(Biobase)
library(impute)
library(edgeR)
library(limma)
suppressMessages(library(GSVA))
library(genefilter)
library(ggplot2)


# Optional

library(knitr)
library(VennDiagram)
flog.threshold(ERROR)
library(RColorBrewer)
library(randomcoloR)

# Set Random Seed for
base::set.seed(42) # Because 42 is the answer to everything in a random universe.
```

```{r functions, results='hide'}

mapGmt2NamedList <- function(gmtFileName=NULL, type='simple') {
  # This function  takes in a GMT file and creates a named list
  if(type == 'Rlist') {
    fieldStart = 2
  }else if(type == 'simple') {
    # means name\tgene1\tgene2\t...
    # if default type is simple set to 2
    fieldStart = 2
  } else if (type == 'broad'){
    # name\tURL\t\gene1\t\gene2\
    # if the default type is broad, set to 3
    fieldStart = 3
  } else {
    # if there default is missing
    # Same as broad but 2nd position is not necessarily URL
    fieldStart = 3
  }
  geneSet <- list()
  if (type=='Rlist') {
    load(file = gmtFileName,verbose = FALSE)
    gmtFile <- gmt
    rm(gmt)
    for (entry in gmtFile) {
      geneSet[[unlist(entry[1])[1]]] <- entry[fieldStart:length(entry)]
    }
  } else {
    geneSet <- list()
    gmtFile <- readLines(con=gmtFileName)
    gmtFile <- strsplit(gmtFile, '\t')
    
    for (entry in gmtFile) {
      entry <- unlist(entry)
      geneSet[[unlist(entry[1])[1]]] <- entry[fieldStart:length(entry)]
    }
  }
  
  rm(gmtFile,entry, gmtFileName,type)
  
  geneSet
}



gradient.hm <- function(genes=geneSet,
                        AssocData=toptable,
                        PathwayData=NULL,
                        Estimate='logFC',
                        const=F,
                        max=NULL,
                        numGenes=50,
                        orderBy="consensus",
                        annotate_col_labels=NULL,
                        annotate_row_labels=NULL,
                        main="Checkerboard",
                        silent=F){
  
  ## Get unique gene symbols from all pathways
  unique.genes <- na.omit(unique(do.call(c, genes)))
  
  ## Create matrix.  One row for each pathway.  One column for each
  ## unique gene.
  mat <- matrix(NA,
                nrow=length(genes),
                ncol=length(unique.genes))
  colnames(mat) <- unique.genes
  rownames(mat) <- names(genes)
  
  ## Fill in values of matrix with Estimates for each gene.  This fills
  ## in a cell even if the gene is not a pathway.
  for(i in rownames(mat)){ # Loop over each pathway
    for(j in colnames(mat)){ # Loop over each gene.
      if (j %in% genes[[i]]) {
        val <- AssocData[AssocData$symbol == j, Estimate]
        if (length(val) == 1) {
          mat[i,j] <- val
        } else {
          mat[i,j] <- NA
        }
      } else
        mat[i,j] <- NA
    }
  }
  
  # remove genes that have no expression values and order by p.value
  mat <- mat[, colSums(is.na(mat)) !=nrow(mat), drop=F]
  
  AssocData <- AssocData[colnames(mat), ]
  
  # order genes by consensus, logfc or p.value
  for (order in orderBy) {
    if (order == "consensus"){
      mat <- mat[, order(colSums(is.na(mat))), drop = F]
    } else {
      if (range(AssocData[[order]])[2] > 1) { # Check if logFC
        mat <- mat[, order(abs(AssocData[[order]]), decreasing = T), drop = F]
      } else {
        mat <- mat[, order(AssocData[[order]]), drop = F]
      }
    }
  }
  
  # trim gene list if too long
  if (dim(mat)[2] > numGenes) {
    mat <- mat[,1:numGenes, drop=F]
  }
  
  # remove pathways that have no genes
  mat <- mat[rowSums(is.na(mat)) !=ncol(mat), , drop=F]
  
  # Check if any genes are left
  if (is.null(mat)) {
    return("No genes pass filter")
  }
  
  ## Order columns of mat by regression coeffiecient.  This makes it so
  ## that all genes in the first pathway will be shown on the left side
  ## of the plot.
  for(i in 1:nrow(mat)){
    mat <- mat[,order(mat[i,]), drop=F]
  }
  
  ## Subset final mat
  mat <- mat[nrow(mat):1, , drop=F]
  
  if(const == T){
    if (is.null(max)) {
      max <- max(abs(mat), na.rm=T)
    }
    seq <- seq(from = -max, to = max, length.out=20)
    colors <- colorRampPalette(c('#1d587c','#2b83ba', '#abdda4','#ffffbf','#fdae61','#d7191c','#891012'))(length(seq))
  } else {
    seq <- seq(floor(min(mat, na.rm=T)), ceiling(max(mat, na.rm=T)), length.out=99)
    colors <- colorRampPalette(c('#2b83ba', '#abdda4','#ffffbf','#fdae61','#d7191c'))(99)
  }
  
  ## Create column annotation if present
  if (!is.null(annotate_col_labels)) {
    annotate_col <- AssocData[colnames(mat), annotate_col_labels, drop=F]
  } else {
    annotate_col <- NULL
  }
  
  ## Create row annotation if present
  if (!is.null(annotate_row_labels) & !is.null(PathwayData)) {
    annotate_row <- PathwayData[rownames(mat), annotate_row_labels, drop=F]
  } else {
    annotate_row <- NULL
  }
  
  
  ## Create plot
  
  require(pheatmap)
  
  pheatmap(mat,
           color=colors,
           breaks=seq,
           annotation_col=annotate_col,
           annotation_row=annotate_row,
           annotation_names_row=F,
           annotation_names_col=F,
           border_color='azure2',
           cluster_cols=FALSE,
           cluster_rows=FALSE,
           cellheight = 12,
           cellwidth = 12,
           main = main,
           labels_row = substr(rownames(mat),1,40),
           fontsize=6,
           silent = silent,
           na_col = "#FFFFFF")
  
}

# Based on code from http://michaeljw.com/blog/post/subchunkify/

#' Generate a sub-chunk to be interpreted by knitr.  The enclosing chunk
#' must have "results='asis'"
#'
#' @param g The output to chunkify (only tested with figures to date)
#' @param ... Additional named arguments to the chunk
#' @return NULL
#' @details The chunk is automatically output to the console.  There is
#'   no need to print/cat its result.
#' @export
subchunkify <- local({
  chunk_count <- 0
  function(g, ...) {
    chunk_count <<- chunk_count + 1
    g_deparsed <-
      paste0(deparse(
        function() {g}
      ),
      collapse = '')
    args <- list(...)
    args <-
      lapply(names(args),
             FUN=function(nm, arglist) {
               current <- arglist[[nm]]
               if (length(current) > 1) {
                 stop("Only scalars are supported by subchunkify")
               } else if (is.character(current) | is.factor(current)) {
                 current <- as.character(current)
                 ret <- paste0('"', gsub('"', '\"', current, fixed=TRUE), '"')
               } else if (is.numeric(current) | is.logical(current)) {
                 ret <- as.character(current)
               } else {
                 stop("Unhandled class in subchunkify argument handling")
               }
               paste0(nm, "=", ret)
             },
             arglist=args)
    args <- paste0(unlist(args), collapse=", ")
    chunk_header <-
      paste(
        paste0("{r sub_chunk_", chunk_count),
        if (nchar(args) > 0) {
          paste(",", args)
        } else {
          NULL
        },
        ", echo=FALSE}")
    
    sub_chunk <- paste0(
      "\n```",chunk_header, "\n",
      "(", 
      g_deparsed
      , ")()\n",
      "```\n")
    cat(knitr::knit(text = knitr::knit_expand(text = sub_chunk), quiet = TRUE))
  }
})


lst_geneSets <- list()

lst_geneSets[['Custom Ingenuity Pathway Analysis']]         <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Custom/Custom_IPA_CP_R.v2.gmt',type = 'broad')
lst_geneSets[['MSigDB Hallmark']]           <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/h.all.v7.2.symbols.gmt',type = 'broad')
lst_geneSets[['MSigDB Canonical Pathways']]  <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c2.cp.biocarta.v7.2.symbols.gmt',type = 'broad')
lst_geneSets[['MSigDB KEGG']]      <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c2.cp.kegg.v7.2.symbols.gmt',type = 'broad')
lst_geneSets[['MSigDB Reactome']]  <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c2.cp.reactome.v7.2.symbols.gmt',type = 'broad')
#lst_geneSets[['MSigDB microRNA Targets']]     <- mapGmt2NamedList(gmtFileName = '../../../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c3.mir.v7.2.symbols.gmt',type = 'broad')
#lst_geneSets[['MSigDB Transcriptional Factor Targets']]       <- mapGmt2NamedList(gmtFileName = '../../../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c3.tft.v7.2.symbols.gmt',type = 'broad')
lst_geneSets[['MSigDB GO Biological Process']]      <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c5.go.bp.v7.2.symbols.gmt',type = 'broad')
lst_geneSets[['MSigDB GO Cellular Component']]       <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c5.go.cc.v7.2.symbols.gmt',type = 'broad')
lst_geneSets[['MSigDB GO Molecular Function']]       <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c5.go.mf.v7.2.symbols.gmt',type = 'broad')
lst_geneSets[['MSigDB Oncogenic Signatures']] <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c6.all.v7.2.symbols.gmt',type = 'broad')
lst_geneSets[['MSigDB Immunologic Signatures']]    <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c7.all.v7.2.symbols.gmt',type = 'broad')

#lst_geneSets$GeoDE_microRNA         <- mapGmt2NamedList(gmtFileName = '../../../../Envs/ENVA_Annotations/GMT_Public/GMT_GeoDE/microRNA.gmt.RData',type = 'Rlist')

lst_geneSets[['Celltype Signature Genesets']]    <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_MsigDB/msigdb_v7.2_GMTs/c8.all.v7.2.symbols.gmt',type = 'broad')

lst_geneSets[['ConsensusPathDB Wikipathways']]   <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Public/GMT_CPDB/CPDB_Wikipathways.gmt',type = 'broad')
lst_geneSets[['Custom Damian Pathways']]         <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Custom/Custom_DamianModules.gmt',type = 'broad')
lst_geneSets[['Custom Nakaya Pathways']]        <- mapGmt2NamedList(gmtFileName = '../../Envs/ENVA_Annotations/GMT_Custom/Custom_Nakaya.gmt',type = 'broad')

colorSets <- function(n=2,colors.used=NULL) {
  if ((n + length(colors.used)) <= 30) {
    colors <- c('#e6194b', '#3cb44b', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', 
      '#800000', '#aaffc3', '#808000', '#ffd8b1', '#000075', '#808080', '#ffbf00', '#e6ccb3', '#8080ff', '#660000', '#006699', '#006600', 
      '#660033', '#336699', '#cccc00', '#2d8659', '#666699', '#666666')
    if (!is.null(colors.used)) {
      colors <- colors[!colors %in% colors.used]
    }
    return(colors[1:n])
  } else {
	  distinctColorPalette(n)
  }
}

```

```{r read_project, results='hide'}

## Read in the existing project

AAA_PROJECT <- readRDS(file = "eset_norm_2022-03-22t14_36_02.RDS")

## Put the structures that are needed in the space.
eset_raw <- AAA_PROJECT$Esets$eset_raw

# End of read project
lst_contrast <- readRDS("save/Preliminary_Analysis_2022-03-29t12_02_00.RDS")

```

# Project Summary
Samples are iPSC Cardiomyocyte (CDI and Axol)

Numbers breakdown:


```{r numbersBreakdown}

table1 <- as.data.frame(table(eset_raw$group))
colnames(table1) <- c("Group","Number in group")
knitr::kable(table1,align = 'c')

table2 <- as.data.frame(table(eset_raw$donor))
colnames(table2) <- c("Group", "Number in group")
knitr::kable(table2, align = 'c')


```

# Summary Tables {.tabset .tabset-pills}

```{r showSummaryTable, results='asis'}

table <- list()
table2 <- list()

for (nameOfSubset in names(lst_contrast$Analysis)) {
  table2[[nameOfSubset]] <- data.frame()
  table[[nameOfSubset]] <- data.frame()
  theContrast <- lst_contrast$Analysis[[nameOfSubset]]
  for (i in colnames(theContrast$contrasts)) {
    # First table (degs)
    table[[nameOfSubset]][i, "# of DEGs with P.value <= 0.05"] <- table(theContrast$DEG$results_full[, paste0(i, "_P.Value")] <= 0.05)[2]
    table[[nameOfSubset]][i, "Adjusted P.value <= 0.10"] <- table(theContrast$DEG$results_full[, paste0(i, "_Adj_P.Value")] <= 0.10)[2]
    
    
    table[[nameOfSubset]][is.na(table[[nameOfSubset]])] <- 0
    
    
    for (theGeneSet in names(lst_geneSets)) {
      table2[[nameOfSubset]][theGeneSet, i] <- table(theContrast$GSVA[[theGeneSet]]$results_full[, paste0(i, "_P.Value")] <= 0.05)[2]
    }
    table2[[nameOfSubset]][is.na(table2[[nameOfSubset]])] <- 0
  }
}


# Pathways
for (nameOfSubset in names(lst_contrast$Analysis)) {
  cat("##", nameOfSubset, "\n\n")
  cat("\n\n#### Differentially Expressed Genes\n\n")
  print(knitr::kable(table[[nameOfSubset]], align = 'c'))
  cat("\n\n")

  cat("\n\n#### Pathways\n\n")
  print(knitr::kable(table2[[nameOfSubset]], align = 'c'))
  cat("\n\n")
}

```

# Differentially Expressed Genes (DEGs) {.tabset .tabset-pills}

```{r plotHeatmapTopGenes.subset, results='asis'}

for (nameOfSubset in names(lst_contrast$Analysis)) {
  cat("\n\n##", nameOfSubset, " {.tabset .tabset-pills}\n\n")
  
  # Setup the object
  theContrast <- lst_contrast$Analysis[[nameOfSubset]]
  
  fullEset <- theContrast$DEG$eset
    
  # Setup colors for each phenotype
  phenos <- names(which(apply(pData(fullEset), 2, function(x) length(na.omit(as.numeric(as.character(x))))) == 0))

  phenos.colors <- sapply(phenos, function(pheno) {
			  colors <- colorSets(length(unique(na.omit(fullEset[[pheno]]))))
			  names(colors) <- unique(na.omit(fullEset[[pheno]]))
			  return(colors)})    

  fullDge       <-  DGEList(counts       =   exprs(fullEset),
                                remove.zeros =   TRUE)
  fullDge <-  calcNormFactors(object = fullDge, method = "TMM")
  
  topExprs <- cpm(fullDge, prior.count = 3, log=T)
  
  topTableFull  <- theContrast$DEG$results_full
  
  # Create Anova heatmap (optional)
  if (F) {
  
  cat("##### Anova Heatmap")
  
  topTableTemp <- subset(topTableFull, topTableFull[, "Anova P.Value"] <= 0.05) 
  
  if (dim(topTableTemp)[1] < 50) {
    n = dim(topTableTemp)[1]
  } else {
    n = 50
  }

  topTableTemp <- topTableTemp[order(topTableTemp[, "Anova P.Value"]), ]

  topExprsTemp <- topExprs[rownames(topTableTemp[1:n,]),]
  
  topExprsTemp <- t(scale(t(topExprsTemp)))
  
  annotation_col <- pData(fullEset)[, c("group"), drop = F]
  
  annotation_row <- data.frame(PValue = topTableTemp[["Anova P.Value"]])
  rownames(annotation_row) <- rownames(topTableTemp)
  
  seq <- seq(-3, 3, length.out = 99)
  
  colors <- colorRampPalette(c('#2b83ba', '#abdda4','#ffffbf','#fdae61','#d7191c'))(length(seq))
  
  p.plot <- pheatmap(topExprsTemp, scale="none", annotation_col = annotation_col, annotation_row = annotation_row, breaks = seq, color = colors, main = "Top 50 DEGs by Anova", cellwidth = 12, cellheight = 12, labels_row = topTableTemp[1:n, "gene_id"], labels_col = fullEset$donor.id, cluster_cols = T, silent = T, annotation_colors = phenos.colors)
  
  subchunkify(grid.draw(p.plot$gtable), 
              fig.width = sum(convertWidth(p.plot$gtable$widths, "inches", valueOnly = T), 1), 
              fig.height = sum(convertHeight(p.plot$gtable$heights, "inches", valueOnly = T), 1), 
              fig.cap = "Anova Heatmap")
  }
  
  cat("##### Per Contrast")
  for(sort.method in c("P.Value", "logFC", "PCC")) {
  
    if (!any(grepl("PCC", colnames(topTableFull))) & sort.method == "PCC") {
      next
    }

    cat("\n\n### Top By ", sort.method, "\n\n")
    gene.list <- c()
    
    # Loop through each contrast
    for (i in colnames(theContrast$contrasts)) {
      # Subset eset according to contrast
      if (grepl("Regressions", nameOfSubset)) {
	tempEset <- fullEset
      } else {
        tempEset <- fullEset[, fullEset$group %in% names(which(theContrast$contrasts[, i] != 0))]
      }
      
      # do not cluster if regression and setup colors
      if (grepl("Regressions", nameOfSubset)) {
        cluster.col = F
        
        # order by variable
        tempEset <- tempEset[, order(as.numeric(tempEset[[i]]))]

        # remove missing values
        tempEset <- tempEset[, !is.na(tempEset[[i]])]

	annotation_col <- data.frame(value = as.numeric(tempEset[[i]]))
        colnames(annotation_col)[1] <- i
        rownames(annotation_col) <- colnames(tempEset)
	  
        annotation_col <- cbind(annotation_col, pData(tempEset)[, c("group"), drop = F])
	annotation_col <- annotation_col[, !duplicated(colnames(annotation_col))]

      } else {
        cluster.col = T
        annotation_col <- pData(fullEset)[, c("group"), drop = F]

      }
      
      # Order by donor.id
      # tempEset <- tempEset[, order(tempEset$donor.id)]
      
      topTableTemp <- subset(topTableFull, topTableFull[, paste0(i, "_P.Value")] <= 0.05) # & topTableFull[, paste0(i, "_notLowCount")])
      
      # Update full list
      gene.list <- c(gene.list, rownames(topTableTemp))
      
      if (dim(topTableTemp)[1] < 50) {
        n = dim(topTableTemp)[1]
      } else {
        n = 50
      }
      
      if (sort.method == "P.Value") {
        topTableTemp <- topTableTemp[order(topTableTemp[, paste0(i, "_", sort.method)]), ]
      } else {
        topTableTemp <- topTableTemp[order(abs(topTableTemp[, paste0(i, "_", sort.method)]), decreasing = T), ]
      }
      
      topExprsTemp <- topExprs[rownames(topTableTemp[1:n,]),]
      topExprsTemp <- topExprsTemp[, sampleNames(tempEset)]
      topExprsTemp <- t(scale(t(topExprsTemp)))
      
      annotation_row <- data.frame(logFC = topTableTemp[[paste0(i, "_logFC")]],
                                   PValue = topTableTemp[[paste0(i, "_P.Value")]])
      rownames(annotation_row) <- rownames(topTableTemp)
      
      seq <- seq(-3, 3,length.out = 99)
      
      colors <- colorRampPalette(c('#2b83ba', '#abdda4','#ffffbf','#fdae61','#d7191c'))(length(seq))
      
      p.plot <- pheatmap(topExprsTemp, scale="none", annotation_col = annotation_col, annotation_row = annotation_row, breaks = seq, color = colors, main = paste("Top 50 DEGs by", sort.method, i), cellwidth = 12, cellheight = 12, labels_row = topTableTemp[1:n, "gene_id"], labels_col = tempEset$donor.id, cluster_cols = cluster.col, silent = T, annotation_colors = phenos.colors)
      
      subchunkify(grid.draw(p.plot$gtable), 
                  fig.width = sum(convertWidth(p.plot$gtable$widths, "inches", valueOnly = T), 1), 
                  fig.height = sum(convertHeight(p.plot$gtable$heights, "inches", valueOnly = T), 1), 
                  fig.cap = paste(i, "Heatmap"))
    }
    
    # Setup datatables for loop
    
    table <- lst_contrast$Analysis[[nameOfSubset]]$DEG$results_full
    table <- table[unique(gene.list), ]
    table <- table[, -c(1: which(colnames(table) == "symbol"))]
    table <- table[, !grepl("notLowCount", colnames(table))]
    
    # Fix Colnames
    colnames(table) <- gsub("_", " ", colnames(table))
    
    # Order columns by regulation, p.value and logFC
    table <- table[, c(grep("Regulation", colnames(table)), grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))]
    
    # Shorten numerical values to 3 significant digits
    table[, c(grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))] <- signif(table[, c(grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))], 3)
    rownames(table) <- paste0('<a href="https://www.genecards.org/cgi-bin/carddisp.pl?gene=', 
                              rownames(table), '&keywords=', rownames(table), '" target="_blank">', rownames(table), '</a>')
    subchunkify(datatable(table, extensions = c('Buttons', 'FixedColumns', 'Scroller'), options = list(
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
      scrollX = TRUE,
      fixedColumns = TRUE,
      deferRender = TRUE,
      scrollY = 500,
      scroller = TRUE
    ),escape = F
    ))
   }
}

```

# Differentially Expressed Pathways (DEPs) {.tabset .tabset-pills}
>Top pathways by P.Value <= 0.05
>Genes were filtered by P.Value <= 0.05

```{r savePathways, results='asis'}

for (nameOfSubset in names(lst_contrast$Analysis)) {
  cat("\n\n##", nameOfSubset, " {.tabset .tabset-pills}\n\n")
  
  # Setup the objects
  theContrast <- lst_contrast$Analysis[[nameOfSubset]]
  
  for (theGeneSet in names(lst_geneSets)) {
    pathway.list <- c()
    first <- TRUE
      
    # If being plotted, create title for the first in the geneSet and plot the anova heatmap
    if (first) {
      cat("\n\n###", theGeneSet, " {.tabset .tabset-pills}\n\n")
      first <- FALSE
    }
    
    # Pull expression values and stats for this geneSet
    fullEset <- theContrast$GSVA[[theGeneSet]]$eset.gsva

    # Setup colors for each phenotype
    phenos <- names(which(apply(pData(fullEset), 2, function(x) length(na.omit(as.numeric(as.character(x))))) == 0))
  
    phenos.colors <- sapply(phenos, function(pheno) {
			  colors <- colorSets(length(unique(na.omit(fullEset[[pheno]]))))
			  names(colors) <- unique(na.omit(fullEset[[pheno]]))
			  return(colors)})    

    topTableFull  <- topTableTemp <- theContrast$GSVA[[theGeneSet]]$results_full
    topExprs <- exprs(fullEset)
   
    anova = F
    
    if (dim(topTableTemp)[1] < 50) {
      n = dim(topTableTemp)[1]
    } else {
      n = 50
    }
    
    if (dim(topTableTemp)[1] == 0) { 
      anova = F
    } else if (dim(topTableTemp)[1] == 1) {
      cluster.row = F
    } else {
      cluster.row = T
    }
    
    if (anova) {
    cat("##### Anova Heatmap")
    
    topTableTemp <- topTableTemp[order(topTableTemp[, "Anova P.Value"]), ]
    
    topExprsTemp <- topExprs[rownames(topTableTemp[1:n,]), ,drop=F]
    
    topExprsTemp <- t(scale(t(topExprsTemp)))
    
    annotation_col <- pData(fullEset)[, c("group"), drop = F]
    
    annotation_row <- data.frame(PValue = topTableTemp[["Anova P.Value"]])
    rownames(annotation_row) <- rownames(topTableTemp)
    
    max <- max(abs(topExprsTemp))
    min <- min(abs(c(quantile(topExprsTemp, .01), quantile(topExprsTemp, .99))))
    
    seq <- seq(-3, 3, length.out = 99)
    
    colors <- colorRampPalette(c('#2b83ba', '#abdda4','#ffffbf','#fdae61','#d7191c'))(length(seq))
    
    p.plot <- pheatmap(topExprsTemp, scale="none", annotation_col = annotation_col, annotation_row = annotation_row, breaks = seq, color = colors, main = "Top 50 DEPs by Anova", cellwidth = 12, cellheight = 12, labels_row = rownames(topTableTemp[1:n, ]), labels_col = fullEset$donor.id, cluster_cols = T, cluster_rows = cluster.row, silent = T, annotation_colors = phenos.colors)
    
    subchunkify(grid.draw(p.plot$gtable), 
              fig.width = sum(convertWidth(p.plot$gtable$widths, "inches", valueOnly = T), 1), 
              fig.height = sum(convertHeight(p.plot$gtable$heights, "inches", valueOnly = T), 1), 
              fig.cap = "Anova Heatmap")
    }

    cat("\n\n##### Per Contrast")
    
    # Loop through each Contrast
    for(sort.method in c("P.Value", "logFC", "PCC")) {
    
      if (!any(grepl("PCC", colnames(topTableFull))) & sort.method == "PCC") {
        next
      }

      cat("\n\n#### Top By ", sort.method, "\n\n")
      gene.list <- c()
      
      for (i in colnames(theContrast$contrasts)) {
        # Subset eset according to contrast
        if (grepl("Regressions", nameOfSubset)) {
          tempEset <- fullEset
        } else {
          tempEset <- fullEset[, fullEset$group %in% names(which(theContrast$contrasts[, i] != 0))]
        }
        
        # do not cluster if regression and setup colors
        if (grepl("Regressions", nameOfSubset)) {
          cluster.col = F
          
          # order by group
          tempEset <- tempEset[, order(as.numeric(tempEset[[i]]))]
        
          # remove missing values
          tempEset <- tempEset[, !is.na(tempEset[[i]])]
        
	  annotation_col <- data.frame(value = as.numeric(tempEset[[i]]))
          colnames(annotation_col)[1] <- i
          rownames(annotation_col) <- colnames(tempEset)

          annotation_col <- cbind(annotation_col, pData(tempEset)[, c("group"), drop = F])
	  annotation_col <- annotation_col[, !duplicated(colnames(annotation_col))]

        } else {
          cluster.col = T
          annotation_col <- pData(fullEset)[, c("group"), drop = F]

        }
        
        topTableTemp <- subset(topTableFull, topTableFull[, paste0(i, "_P.Value")] <= 0.05) # & topTableFull[, paste0(i, "_notLowCount")])
        pathway.list <- c(pathway.list, rownames(topTableTemp))
        
        if (dim(topTableTemp)[1] < 50) {
          n = dim(topTableTemp)[1]
        } else {
          n = 50
        }
        
        if (dim(topTableTemp)[1] == 0) {
          next
        } else if (dim(topTableTemp)[1] == 1) {
          cluster.row = F
        } else {
          cluster.row = T
        }
        
        if (sort.method == "P.Value") {
          topTableTemp <- topTableTemp[order(topTableTemp[, paste0(i, "_", sort.method)]), ]
        } else {
          topTableTemp <- topTableTemp[order(abs(topTableTemp[, paste0(i, "_", sort.method)]), decreasing = T), ]
        }
        
        # Setup heatmap values
        topExprsTemp <- topExprs[rownames(topTableTemp[1:n,]), ,drop=F]
        
        # Order by Regression var
        #tempEset <- tempEset[, order(tempEset[[i]]), drop = F]
        topExprsTemp <- topExprsTemp[, sampleNames(tempEset), drop = F]
        
        topExprsTemp <- t(scale(t(topExprsTemp)))
        
        annotation_row <- data.frame(logFC = topTableTemp[[paste0(i, "_logFC")]],
                                     PValue = topTableTemp[[paste0(i, "_P.Value")]])
        rownames(annotation_row) <- rownames(topTableTemp)
        
        seq <- seq(-3, 3, length.out = 99)
        
        colors <- colorRampPalette(c('#2b83ba', '#abdda4','#ffffbf','#fdae61','#d7191c'))(length(seq))
      
        p.plot <- pheatmap::pheatmap(topExprsTemp, scale="none", annotation_col = annotation_col, annotation_row = annotation_row, breaks = seq, color = colors, main = paste("Top 50 DEPs by", sort.method, i), cellwidth = 12, cellheight = 12, fontsize = 6, labels_row = topTableTemp[1:n, "gene_id"], labels_col = tempEset$donor.id, cluster_cols = cluster.col, cluster_rows = cluster.row, silent = T, annotation_colors = phenos.colors)
        
        fig.width = sum(convertWidth(p.plot$gtable$widths, "inches", valueOnly = T), 1)
        fig.height = sum(convertHeight(p.plot$gtable$heights, "inches", valueOnly = T), 1)
        
        subchunkify(grid.draw(p.plot$gtable), 
                  fig.width = fig.width, 
                  fig.height = ifelse(fig.height < 4, fig.height + (4 - fig.height), fig.height), 
                  fig.cap = paste(i, "Heatmap"),
                  fig.retina = 2)
        
        #### Create the checkerboard
        topGenes  <- theContrast$DEG$results_full
        topGenes <- subset(topGenes, topGenes[, paste0(i, "_P.Value")] <= 0.05)
        
        geneList <- lst_geneSets[[theGeneSet]][rownames(topTableTemp[1:n,])]
        if (length(geneList) < 20) {
          geneList <- geneList[1:length(geneList)]
        } else {
          geneList <- geneList[1:20]
        }
        
        if (sum(unique(unlist(geneList)) %in% rownames(topGenes)) < 2) {
          next
        }
        
        p.plot <- gradient.hm(genes = geneList
                    , AssocData = topGenes
                    , PathwayData = topTableFull
                    , Estimate = paste0(i, "_logFC")
                    , orderBy = c(paste0(i, "_", sort.method), "consensus")
                    , const = T
                    , annotate_col = paste0(i, "_P.Value")
                    , annotate_row = c(paste0(i, "_logFC"))
                    , main = paste(i, theGeneSet, "By", sort.method)
                    , silent = T)
        
        fig.width = sum(convertWidth(p.plot$gtable$widths, "inches", valueOnly = T), 1)
        fig.height = sum(convertHeight(p.plot$gtable$heights, "inches", valueOnly = T), 1)
        
        subchunkify(grid.draw(p.plot$gtable), 
                  fig.width = fig.width, 
                  fig.height = ifelse(fig.height < 3, fig.height + (3-fig.height), fig.height), 
                  fig.cap = paste(i, "Checkerboard"),
                  fig.retina = 2)
        
      }
      
      rm(table)
      
      table <- theContrast$GSVA[[theGeneSet]]$results_full
      table <- table[unique(pathway.list), ]
      #table <- table[, -c(1: which(colnames(table) == "symbol"))]
      
      # Fix Colnames
      colnames(table) <- gsub("_", " ", colnames(table))
      
      # Order columns by regulation, p.value and logFC
      table <- table[, c(grep("Regulation", colnames(table)), grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))]
      
      # Shorten numerical values to 3 significant digits
      table[, c(grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))] <- signif(table[, c(grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))], 3)
      
      # # Fix Rownames
      # table <- cbind("Pathway" = gsub("_", " ", rownames(table)), table)
      # rownames(table) <- NULL

      subchunkify(datatable(table, extensions = c('Buttons', 'FixedColumns', 'Scroller'),
                            options = list(
                              autowidth = TRUE,
                              dom = 'Bfrtip',
                              buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                              deferRender = TRUE,
                              fixedColumns = TRUE,
                              scrollY = 500,
                              scrollX = 200,
                              scroller = TRUE,
                              columnDefs = list(list(
                                targets = 0,
                                render = JS(
                                  "function(data, type, row, meta) {",
                                  "return type === 'display' && data.length > 20 ?",
                                  "'<span title=\"' + data + '\">' + data.substr(0, 20) + '...</span>' : data; ",
                                  "}")
                              ))
                            )
      ))
    }
  }
}


```


# Session Info

```{r session_info, results='markup'}
Hmisc::markupSpecs$html$session()
```

# Reference Tables {.tabset .tabset-pills}

## Full Gene Table

```{r setup Final Gene Table}
table <- c()
for(nameOfSubset in names(lst_contrast$Analysis)) { 
  
  tempTable <- lst_contrast$Analysis[[nameOfSubset]]$DEG$results_full
  
  # Fix Colnames
  colnames(tempTable) <- gsub("_", " ", colnames(tempTable))
  colnames(tempTable) <- paste(nameOfSubset, colnames(tempTable))
  
  if (is.null(dim(table))) {
    table <- tempTable
  } else {
    ids <- c(rownames(table, tempTable))
    table <- cbind(table[ids, ], tempTable[ids, ])
    rownames(table) <- ids
  }
}

# Order columns by regulation, p.value and logFC
table <- table[, c(grep("Regulation", colnames(table)), grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))]

# Shorten numerical values to 3 significant digits
table[, c(grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))] <- signif(table[, c(grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))], 3)

datatable(table, extensions = c('Buttons', 'FixedColumns', 'Scroller'), options = list(
  dom = 'Bfrtip',
  buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
  scrollX = TRUE,
  fixedColumns = TRUE,
  deferRender = TRUE,
  scrollY = 500,
  scroller = TRUE
)
)

```

## Full Pathway Table

```{r setup Final pathway table, eval = F}


# Generate a searchable table instead of excel worksheets
# Concatenate all of the test statistics

for(nameOfSubset in names(lst_contrast$Analysis)) { 
  tempTableSet <- c()
  for(theGeneSet in names(lst_geneSets)) {
    tempTable <- lst_contrast$Analysis[[nameOfSubset]]$GSVA[[theGeneSet]]$results_full
    
    # Fix Colnames
    colnames(tempTable) <- gsub("_", " ", colnames(tempTable))
    colnames(tempTable) <- paste(nameOfSubset, colnames(tempTable))
    
    if (is.null(dim(tempTableSet))) {
      tempTableSet <- cbind("Pathway" = rownames(tempTable), "Pathway DB" = rep(theGeneSet, dim(tempTable)[1]), tempTable, "uniqueID" = paste(theGeneSet, rownames(tempTable), sep = "|"))
    } else {
      tempTableSet <- rbind(tempTableSet, cbind("Pathway" = rownames(tempTable), "Pathway DB" = rep(theGeneSet, dim(tempTable)[1]), tempTable, "uniqueID" = paste(theGeneSet, rownames(tempTable), sep = "|")))
    }
  }
  if (is.null(dim(table))) {
    table <- tempTableSet
  } else {
    ids <- unique(c(as.character(table$uniqueID), as.character(tempTableSet$uniqueID)))
    table <- cbind(table[match(ids, table$uniqueID), ], tempTableSet[match(ids, tempTableSet$uniqueID), ])
  }
}

# Order columns by regulation, p.value and logFC
table <- table[, c(1:2, grep("Regulation", colnames(table)), grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))]

# Shorten numerical values to 3 significant digits
table[, c(grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))] <- signif(table[, c(grep("P.Value", colnames(table)), grep("logFC", colnames(table)), grep("PCC", colnames(table)))], 3)

# # Fix Rownames
# table <- cbind("Pathway" = gsub("_", " ", rownames(table)), table)
# rownames(table) <- NULL

datatable(table, rownames = F, extensions = c('Buttons', 'FixedColumns', 'Scroller'),      
          options = list(
            autowidth = TRUE,
            dom = 'Bfrtip',
            buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
            deferRender = TRUE,
            fixedColumns = TRUE,
            scrollY = 500,
            scrollX = 200,
            scroller = TRUE,
            columnDefs = list(list(
              targets = 0,
              render = JS(
                "function(data, type, row, meta) {",
                "return type === 'display' && data.length > 20 ?",
                "'<span title=\"' + data + '\">' + data.substr(0, 20) + '...</span>' : data; ",
                "}")
            ))
          )
)

```


```{r Session Info, results='markup'}

Hmisc::markupSpecs$html$session()

```
